---
title: "1-Simulate datasets"
author: "Siyuan Ma"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
---
# Overview
- Create table of simulation parameter combinations.
1. Homogeneous studies.
2. Studies with heterogeneous cohort characteristics (i.e. eT distribution).
3. Studies with different design (i.e., censoring distribution).
- Simulate datasets.
```{r setup, echo=FALSE} 
knitr::opts_knit$set(root.dir = normalizePath(".."))
```
```{r setup2, echo=FALSE}
rm(list = ls())
for(i.function in list.files("functions/", full.names = TRUE)) {
  source(i.function)
}
dir_output <- "results/1-simulate_datasets/"
dir.create(dir_output, recursive = TRUE, showWarnings = FALSE)
library(magrittr)
```
# Create table of simulation parameter combinations.
## Homogeneous studies
```{r simulation parameters scenario I}
# number of simulation runs
rep <- 1:500
# total sample size
nSamples <- c(200)
# number of studies
nStudies <- c(2)
# missingness of studies
missingness <- tibble::tibble(missingness = list(c(FALSE, TRUE)))
# distribution of error terms
dist_eT <- tibble::tibble(dist_eT = list(c("normal", "normal")))
sd_eT <- tibble::tibble(sd_eT = list(c(1, 1)))
# percentage censoring
percCens <- tibble::tibble(percCens = list(c(0.1, 0.1),
                                           c(0.4, 0.4)))
# true coefficients
b <- tibble::tibble(b = list(c(1, 1, 0),
                             c(1, 1, 1),
                             c(1, 1, 2)))
a <- tibble::tibble(a = list(matrix(c(1, 1), nrow = 2, ncol = 1)))
tb_sim_I <- tidyr::crossing(
  nSamples,
  nStudies,
  missingness,
  dist_eT,
  sd_eT,
  percCens,
  b,
  a
) %>%
  dplyr::mutate(i_setup = 1:n()) %>% 
  tidyr::crossing(rep) %>% 
  dplyr::mutate(study = nStudies %>% 
                  purrr::map2(nSamples, function(nStudies, nSamples) 
                    rep(1:nStudies, each = nSamples / nStudies)),
                seed = 1:n())
simulation_sanitycheck(tb_sim_I)
```
## Studies with heterngeneous eT distributions
```{r simulation parameters scenario II}
# number of simulation runs
rep <- 1:500
# total sample size
nSamples <- c(400)
# number of studies
nStudies <- c(4)
# missingness of studies
missingness <- tibble::tibble(missingness = list(c(FALSE, FALSE, TRUE, TRUE)))
# distribution of error terms
tb_eT <- tibble::tibble(dist_eT = list(c("normal", "evd", "normal", "evd"),
                                       c("normal", "normal", "evd", "evd"),
                                       c("normal", "normal", "normal", "normal")),
                        sd_eT = list(c(1, 1, 1, 1),
                                     c(1, 1, 1, 1),
                                     c(1, 4, 2, 3)))
# percentage censoring
percCens <- tibble::tibble(percCens = list(c(0.1, 0.1, 0.1, 0.1)))
# true coefficients
b <- tibble::tibble(b = list(c(1, 1, 1)))
a <- tibble::tibble(a = list(matrix(c(1, 1), nrow = 2, ncol = 1)))
tb_sim_II <- tidyr::crossing(
  nSamples,
  nStudies,
  missingness,
  tb_eT,
  percCens,
  b,
  a,
) %>%
  dplyr::mutate(i_setup = 1:n()) %>% 
  tidyr::crossing(rep) %>% 
  dplyr::mutate(study = nStudies %>% 
                  purrr::map2(nSamples, function(nStudies, nSamples) 
                    rep(1:nStudies, each = nSamples / nStudies)),
                seed = 1:n())
simulation_sanitycheck(tb_sim_II)
```
## Studies with heterngeneous ceonsoring distributions
```{r simulation parameters scenario III}
# number of simulation runs
rep <- 1:500
# total sample size
nSamples <- c(400)
# number of studies
nStudies <- c(4)
# missingness of studies
missingness <- tibble::tibble(missingness = list(c(FALSE, FALSE, TRUE, TRUE)))
# distribution of error terms
dist_eT <- tibble::tibble(dist_eT = list(c("normal", "normal", "normal", "normal")))
sd_eT <- tibble::tibble(sd_eT = list(c(1, 1, 1, 1)))
# percentage censoring
percCens <- tibble::tibble(percCens = list(c(0.1, 0.4, 0.1, 0.4),
                                           c(0.1, 0.1, 0.4, 0.4),
                                           c(0.1, 0.4, 0.2, 0.3)))
# true coefficients
b <- tibble::tibble(b = list(c(1, 1, 1)))
a <- tibble::tibble(a = list(matrix(c(1, 1), nrow = 2, ncol = 1)))
tb_sim_III <- tidyr::crossing(
  nSamples,
  nStudies,
  missingness,
  dist_eT,
  sd_eT,
  percCens,
  b,
  a,
) %>%
  dplyr::mutate(i_setup = 1:n()) %>% 
  tidyr::crossing(rep) %>% 
  dplyr::mutate(study = nStudies %>% 
                  purrr::map2(nSamples, function(nStudies, nSamples) 
                    rep(1:nStudies, each = nSamples / nStudies)),
                seed = 1:n())
simulation_sanitycheck(tb_sim_III)
```

# Simulate datasets.
```{r simulate datasets}
tb_sim_I <- tb_sim_I %>% 
  dplyr::group_by(1:n()) %>% 
  dplyr::mutate(dataset = simulation(b = b[[1]],
                                     a = a[[1]],
                                     study = study[[1]],
                                     missingness = missingness[[1]],
                                     dist_eT = dist_eT[[1]],
                                     sd_eT = sd_eT[[1]],
                                     percCens = percCens[[1]],
                                     seed = seed) %>% 
                  list()) %>% 
  dplyr::ungroup() %>% 
  dplyr::select(-`1:n()`)
tb_sim_II <- tb_sim_II %>% 
  dplyr::group_by(1:n()) %>% 
  dplyr::mutate(dataset = simulation(b = b[[1]],
                                     a = a[[1]],
                                     study = study[[1]],
                                     missingness = missingness[[1]],
                                     dist_eT = dist_eT[[1]],
                                     sd_eT = sd_eT[[1]],
                                     percCens = percCens[[1]],
                                     seed = seed) %>% 
                  list()) %>% 
  dplyr::ungroup() %>% 
  dplyr::select(-`1:n()`)
tb_sim_III <- tb_sim_III %>% 
  dplyr::group_by(1:n()) %>% 
  dplyr::mutate(dataset = simulation(b = b[[1]],
                                     a = a[[1]],
                                     study = study[[1]],
                                     missingness = missingness[[1]],
                                     dist_eT = dist_eT[[1]],
                                     sd_eT = sd_eT[[1]],
                                     percCens = percCens[[1]],
                                     seed = seed) %>% 
                  list()) %>% 
  dplyr::ungroup() %>% 
  dplyr::select(-`1:n()`)
save(tb_sim_I, file = paste0(dir_output, "scenarioI.RData"))
save(tb_sim_II, file = paste0(dir_output, "scenarioII.RData"))
save(tb_sim_III, file = paste0(dir_output, "scenarioIII.RData"))
```