---
title: "3-Evaluate estimators"
author: "Siyuan Ma"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
---
# Overview
- Evaluate performance of the five estimators x three scenarios.
- mean
- empirical variance
- bias
- MSE
- coverage probability
```{r setup, echo=FALSE} 
knitr::opts_knit$set(root.dir = normalizePath(".."))
```
```{r setup2, echo=FALSE}
rm(list = ls())
for(i.function in list.files("functions/", full.names = TRUE)) {
  source(i.function)
}
dir_output <- "results/3-evaluate_estimators/"
dir.create(dir_output, recursive = TRUE, showWarnings = FALSE)
library(magrittr)
```
```{r load datasets}
load("results/1-simulate_datasets/scenarioI.RData")
load("results/1-simulate_datasets/scenarioII.RData")
load("results/1-simulate_datasets/scenarioIII.RData")
load("results/2-run_estimators/fits.RData")
tb_sim <- rbind(tb_sim_I %>% dplyr::mutate(scenario = 1),
                tb_sim_II %>% dplyr::mutate(scenario = 2),
                tb_sim_III %>% dplyr::mutate(scenario = 3))
```
```{r evaluate estimators}
tb_eval <- tb_sim %>% 
  dplyr::mutate(results = l_results) %>% 
  dplyr::group_by(i_setup, scenario) %>% 
  dplyr::summarise(obs = list(evaluate(truth = b[[1]],
                                       l_results = results %>% purrr::map("obs"))),
                   full = list(evaluate(truth = b[[1]],
                                        l_results = results %>% purrr::map("full"))),
                   opt = list(evaluate(truth = b[[1]],
                                       l_results = results %>% purrr::map("opt"))),
                   fib = list(evaluate(truth = b[[1]],
                                       l_results = results %>% purrr::map("fib"))),
                   mi = list(evaluate(truth = b[[1]],
                                      l_results = results %>% purrr::map("mi")))) %>% 
  dplyr::ungroup() %>% 
  dplyr::group_by(1:n()) %>% 
  dplyr::mutate(result_summarised = rbind(obs[[1]] %>% dplyr::mutate(estimator = "obs"),
                                          full[[1]] %>% dplyr::mutate(estimator = "full"),
                                          opt[[1]] %>% dplyr::mutate(estimator = "opt"),
                                          fib[[1]] %>% dplyr::mutate(estimator = "fib"),
                                          mi[[1]] %>% dplyr::mutate(estimator = "mi")) %>% 
                  dplyr::mutate(i_setup = i_setup, scenario = scenario) %>% 
                  list()) %>% 
  magrittr::extract2("result_summarised") %>% 
  purrr::reduce(rbind) %>% 
  dplyr::mutate(estimator = factor(estimator, levels = c("obs", "full", "opt", "fib", "mi")))
tb_eval <- tb_eval %>% 
  dplyr::left_join(tb_sim %>% dplyr::filter(rep == 1), by = c("i_setup", "scenario")) %>% 
  dplyr::mutate(variable = variable %>% 
                  dplyr::recode_factor(`1` = "M",
                                       `2` = "Z"))
```
# Homogeneous studies
```{r homogeneous}
tb_print <- tb_eval %>% 
  dplyr::filter(scenario == 1) %>% 
  dplyr::mutate(b_Z = b %>% purrr::map_dbl(2), 
                censoring = percCens %>% purrr::map_dbl(unique),
                setup = paste0("b_Z=", b_Z, ", ", censoring, " censoring")) %>% 
  dplyr::arrange(censoring, b_Z, variable, estimator) %>% 
  dplyr::select(setup, variable, estimator, mean, sd, bias, MSE, coverage)
```
Performance on b_M (i.e., coefficient for the mediator) varies between estimators:
```{r homogeneous print b_M}
tb_print %>% 
  dplyr::filter(variable == "M") %>% 
  print()
```
Whereas for the missing covariates b_Z not so much:
```{r homogeneous print b_M}
tb_print %>% 
  dplyr::filter(variable == "Z") %>% 
  print()
```
# Studies with heterngeneous eT distributions
```{r heterogeneous e_T}
tb_print <- tb_eval %>% 
  dplyr::filter(scenario == 2) %>%
  dplyr::mutate(dist_eT = dist_eT %>% purrr::map_chr(paste, collapse = ","), 
                sd_eT = sd_eT %>% purrr::map_chr(paste, collapse = ","),
                setup = paste0("distribution ", dist_eT, ", sd(e_T) ", sd_eT)) %>% 
  dplyr::arrange(dist_eT, sd_eT, variable, estimator) %>% 
  dplyr::select(setup, variable, estimator, mean, sd, bias, MSE, coverage)
tb_print %>% 
  dplyr::filter(variable == "M") %>% 
  print()
```
# Studies with heterngeneous censoring distribution
```{r heterogeneous percCens}
tb_print <- tb_eval %>% 
  dplyr::filter(scenario == 3) %>%
  dplyr::mutate(censoring = percCens %>% purrr::map_chr(paste, collapse = ","), 
                setup = paste0("censoring ", censoring)) %>% 
  dplyr::arrange(censoring, variable, estimator) %>% 
  dplyr::select(setup, variable, estimator, mean, sd, bias, MSE, coverage)
tb_print %>% 
  dplyr::filter(variable == "M") %>% 
  print()
```